Poincar\'e--Friedrichs constants over domains with shellable triangulations

### 2024-06-28 main.tex
some clean up

We discuss the Sobolev spaces of vector calculus and their transformation behavior. 
Let $\Omega \subseteq \bbR^{3}$ be a bounded open set. 
First, recall that $L^{p}(\Omega)$ is the space of scalar-valued $p$-integrable functions defined on $\Omega$
and that $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ for vector-valued functions with each component in $L^{p}(\Omega)$. 
In the three-dimensional setting, we are particularly interested in the Sobolev vector analysis. 
$W^p(\grad,\Omega) := W^{1,p}(\Omega)$ is the space of scalar-valued $L^{p}(\Omega)$ functions with weak gradients in $\bfL^{p}(\Omega)$, and we have 
\[
    W^p(\grad,\Omega) = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
\]
The space $W^{p}(\curl,\Omega)$ of vector-valued $\bfL^{p}(\Omega)$ functions with weak curls in $\bfL^{p}(\Omega)$
and the space of vector-valued $\bfL^{p}(\Omega)$ functions with weak divergences in $L^{p}(\Omega)$ are defined as 
\[
    \bfW^{p}(\curl,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \curl \bfu \in \bfL^{p}(\Omega) \},
    \quad 
    \bfW^{p}(\divergence,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \divergence \bfu \in L^{p}(\Omega) \}.
\]
We are interested in transformations of these Sobolev tensor fields from one domain onto the other. 
Suppose that $\Omega, \Omega' \subset \mathbb{R}^3$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
The gradient-, curl-, divergence-conforming Piola transformations are the mappings 
$\phi^{\grad}: L^{p}(\Omega') \to L^{p}(\Omega)$,
$\phi^{\curl}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$, 
$\phi^{\divergence}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$,
and
$\phi^{\rm b}: L^{p}(\Omega') \to L^{p}(\Omega)$. 
They are defined by 
\begin{subequations}\label{eq_definition_piola}
\begin{align}
    \phi^{\grad}(v) &= v \circ \phi, \\
    \phi^{\curl}(\bfw) &= (\Jacobian\phi^{T} \bfw) \circ \phi, \\
    \phi^{\divergence}(\bfw) &= \left( \det(\Jacobian\phi) \Jacobian\phi^{-1} \bfw \right)  \circ \phi, \\
    \phi^{\rm b}(v) &= \left( \det\Jacobian\phi^{-1} v \right)  \circ \phi,
\end{align}
\end{subequations}
where $v \in L^{p}(\Omega')$ and $\bfw \in \bfL^{p}(\Omega')$, 
and where $\Jacobian\phi$ is the Jacobian matrix of $\phi$. 
These mappings are invertible. 
%
% TODO: bounds
%
We use the commutativity relations 
\begin{subequations}\label{eq_piola_commute}
\begin{align}
    \grad (\phi^{\grad} (v)) &= \phi^{\curl}(\grad v), 
    \\
    \curl (\phi^{\curl}(\bfv)) &= \phi^{\divergence}(\curl \bfv) 
    % TODO: watch the sign 
    \\
    \divergence (\phi^{\divergence}(\bfw)) &= \phi^{\rm b}(\divergence \bfw)
\end{align}
\end{subequations}
for $v \in W^{p}(\grad,\Omega)$, $\bfv \in W^{p}(\curl,\Omega)$, and $\bfw \in W^{p}(\divergence,\Omega)$.  
We summarize this as a commuting diagram:
\begin{align*}
    \begin{CD}
        W^{1,p}(\grad,\Omega') @>{\grad}>> \bfW^{p}(\curl,\Omega') @>{\curl}>> \bfW^{p}(\divergence,\Omega') @>{\divergence}>> L^{p}(\Omega')
        \\
        @VV{\phi^{\grad}}V 
        @VV{\phi^{\curl}}V 
        @VV{\phi^{\divergence}}V 
        @VV{\phi^{\rm b}}V 
        \\
        W^{1,p}(\grad,\Omega ) @>{\grad}>> \bfW^{p}(\curl,\Omega ) @>{\curl}>> \bfW^{p}(\divergence,\Omega ) @>{\divergence}>> L^{p}(\Omega )
    \end{CD}
\end{align*}

\begin{remark}
    Our notion of Piola transform along a mapping $\phi : \Omega \rightarrow \Omega'$ goes in the opposite direction of the mapping:
    scalar and vector fields over $\Omega'$ are transformed into scalar and vector fields over $\Omega'$.
    This is in accordance with the notion of pullback, that we will review shortly. 
    One advantage of that definition is that it also makes sense whenever the transformation is not a diffeomorphism. 
\end{remark}

@@

We also assume the perspective of exterior calculus. 
We give a brief introduction into this topic. 
Let $V$ be a real vector space. 
Given an integer $k \geq 0$, We let $\Alt^{k}(V)$ denote the space of scalar-valued antisymmetric $k$-linear forms over $V$. 
Recall that any $k$-linear scalar-valued form $u$ over $V$ is called antisymmetric
if 
\[ 
    u( v_{\pi(1)}, v_{\pi(2)}, \ldots, v_{\pi(k)} ) 
    = 
    \text{sign}(\pi) 
    u( v_1, v_2, \ldots, v_k ) 
\]
for any $v_1, v_2, \dots, v_k \in V$ and any permutation $\pi$ of the indices \(\{1, 2, \ldots, k\}\). 
By definition, $\Alt^{1}(V)$ is just the dual space of $V$, and $\Alt^{0}(V)$ is the space of real numbers. 
Formally, we define $\Alt^{k}(V)$ to be the zero vector space when $k < 0$.

The wedge product (or exterior product) of alternating multilinear forms is a fundamental operation in exterior algebra. 
Given two alternating multilinear forms \( u_1 \in \Alt^{k}(V) \) and \( u_2 \in \Alt^{l}(V) \), 
their wedge product \( u_1 \wedge u_2 \) is a member of $\Alt^{k+l}(V)$
defined by the formula 
\[
    (u_1 \wedge u_2)(v_1, v_2, \ldots, v_{k+l}) 
    = 
    \frac{1}{k! l!} 
    \sum_{ \sigma \in S_{k+l}} 
    \signum(\sigma) 
    u_1( v_{\sigma(  1)}, \ldots, v_{\sigma(k  )} ) 
    u_2( v_{\sigma(k+1)}, \ldots, v_{\sigma(k+l)} ),
\]
for any \( v_1, v_2, \ldots, v_{k+l} \in V \).
Here, \( S_{k+l} \) denotes the symmetric group on \( k+l \) elements, and \( \signum(\sigma) \) is the sign of the permutation \( \sigma \). The exterior product is bilinear and associative, and satisfies 
\begin{align*}
    u_1 \wedge u_2 = (-1)^{kl} u_2 \wedge u_1,
    \quad 
    u_1 \in \Alt^{k}(V),
    \quad 
    u_2 \in \Alt^{l}(V).
\end{align*}
We are interested only in the special case $V = \bbR^{n}$ of alternating forms over the $n$-dimensional Euclidean space. 
Here, it is customary to identify $\Alt^{k}(V)$ with the space of antisymmetric tensors in $k$ indices. 
Moreover, this particular setting comes with a canonical basis. 
We let \(\{\cartanx^1, \cartanx^2, \ldots, \cartanx^{n}\}\) be the canonical basis dual to the standard unit vectors.
This is a canonical basis of $\Alt^{1}(\bbR^{n})$. 
To define a canonical basis of $\Alt^{k}(\bbR^{n})$, 
we first introduce $\Sigma(k,n)$, the set of strictly ascending mappings $\sigma : \{1,\dots,k\} \rightarrow \{1,\dots,n\}$,
and introduce the basic $k$-alternators 
\begin{align*}
    \cartanx^{\sigma(1)} \wedge \dots \wedge \cartanx^{\sigma(k)}, \quad \sigma \in \Sigma(k,n). 
\end{align*}
These define a basis of $\Alt^{k}(\bbR^{n})$.
Note that 
$\dim \Alt^{k}(\bbR^{n}) = \binom{n}{k}$. 
In particular, $\Alt^{k}(\bbR^{n})$ is the zero vector space whenever $k > n$.
% The dimension of the space of \(k\)-forms, denoted \(\bigwedge^k V\), is given by the binomial coefficient \(\binom{n}{k}\).

@@

We write $C^{\infty}\Alt^{k}(\Omega)$ for the space of smooth differential $k$-forms over $\Omega \subseteq \bbR^{n}$,
which is the vector space of smooth mappings from $\Omega$ into $\Alt^{k}(\bbR^{n})$.
The exterior derivative \( d \) is an operator that takes a \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) 
to a \((k+1)\)-form \( d\omega \in C^{\infty}\Alt^{k+1}(\Omega) \). 
Every \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) can be written 
\[
    \omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \omega_{\sigma} \, 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)},
\]
where \( \omega_{\sigma} : \Omega \rightarrow \bbR \) are smooth functions.
The exterior derivative \( d\omega \) is defined by
\[
    d\omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \sum_{j=1}^{n} 
    \frac{\partial \omega_{\sigma}}{\partial x^j} 
    \, \cartanx^j \wedge 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)}.
\]
The exterior derivative is linear and nilpotent, which means 
\( d(d\omega) = 0 \) for any \( \omega \in C^{\infty}\Alt^{k}(\Omega) \).
Moreover, it satisfies the Leibniz rule:
\[ 
    d(\omega \wedge \eta) 
    = 
    d\omega \wedge \eta + (-1)^{k} \omega \wedge d\eta, 
    \qquad \omega \in C^{\infty}\Alt^{k}(\Omega), 
    \qquad \eta \in C^{\infty}\Alt^{l}(\Omega).
\]
Let u now turn our attention to Sobolev spaces of differential forms. 
Since $\Alt^{k}(\Omega)$ is a normed space, we can consider the pointwise norms of differential $k$-forms. 
% First, we notice that the % TODO: introduce a norm 
We let $L^{p}\Alt^{k}(\Omega)$ be the space of differential $k$-forms over $\Omega$ with locally integrable coefficients 
such that its pointwise norm is $p$-integrable. 
The exterior derivative is defined in the sense of distributions and we introduce 
\[
    W^p\Alt^{k}(\Omega) = \{ u \in L^{p}\Alt^{k}(\Omega) \suchthat \cartan v \in L^{p}\Alt^{k+1}(\Omega) \}
\]
We are interested in transformations of these Sobolev tensor fields from one domain onto the other. 
Suppose that $\Omega,\Omega' \subset \bbR^{n}$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
The pullback of $u \in W^p\Alt^{k}(\Omega')$ along $\phi$ is the differential form 
\[ 
    \phi^{\ast}u_{|x}( v_{1}, v_{2}, \ldots, v_{k} ) 
    = 
    u_{|\phi(x)}( \Jacobian\phi_{|x} \cdot v_1, \Jacobian\phi_{|x} \cdot v_2, \ldots, \Jacobian\phi_{|x} \cdot v_k ) 
\]
for any $v_1, v_2, \dots, v_k \in \bbR^{n}$ and any $x \in \Omega$. One can show that $\phi^{\ast}u \in W^p\Alt^{k}(\Omega)$.

\begin{remark}
    In three dimensions, 
    the calculus of differential forms can be translated into classical vector calculus. 
    This is expressed formally as the commuting diagram 
    \[
    \begin{CD}
        C^\infty\Alt^0(\Omega) @>{d}>> C^\infty\Alt^1(\Omega) @>{d}>> C^\infty\Alt^2(\Omega) @>{d}>> C^\infty\Alt^3(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        @VV{\varpi^{3}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{3} @>{\curl}>> C^\infty(\Omega)^{3} @>{\divergence}>> C^\infty(\Omega)
    \end{CD}
    \]
    where $\varpi^{0}$ and $\varpi^{3}$ are the identity mappings and where 
    \begin{align*}
     \varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} + u_3 \cartanx^{3} \right) 
     &= 
     \left( u_1, u_2, u_3 \right), 
     \\
     \varpi^{2}\left( u_{12} \cartanx^{1} \wedge \cartanx^{2} + u_{13} \cartanx^{1} \wedge \cartanx^{3} + u_{23} \cartanx^{2} \wedge \cartanx^{3} \right) 
     &= 
     \left( u_{23}, - u_{13}, u_{12} \right).   
    \end{align*}
    In two dimensions, 
    the calculus of differential forms can be translated into 2D vector calculus in two different ways. 
    To the authors' best knowledge, neither convention is dominant in the literature.
    We summarize the situation in the following commuting diagram: 
    \[
    \begin{CD}
        C^\infty(\Omega) @>{\curl}>> C^\infty(\Omega)^{2} @>{\divergence}>> C^\infty(\Omega)
        \\
        @AA{\varkappa^{0}}A 
        @AA{\varkappa^{1}}A 
        @AA{\varkappa^{2}}A 
        \\
        C^\infty\Alt^0(\Omega) @>{d}>> C^\infty\Alt^1(\Omega) @>{d}>> C^\infty\Alt^2(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{2} @>{\rot}>> C^\infty(\Omega)
    \end{CD}
    \]
    Here, $\varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} \right) = \left( u_1, u_2 \right)$ is the lower middle isomorphism. We introduce the rotation operator $J(x,y) = (y,-x)$ and define $\varkappa = J \varpi$ and $\rot = \divergence J$.
    The other vertical arrows are the identity. 
    The utility of exterior calculus is that the operators of vector calculus can be translated into a common framework that does not depend on the dimension.
\end{remark}

@@

    T_{\ell,a} u(x) 

@@

    \cartan T_{\ell,a} u(x) 

@@

    k

@@

    T_{\ell+1,a} \cartan u(x)

@@

In particular, if $\cartan u = 0$, then $\cartan T_{\ell,a} u = u$.

>>>>

We discuss the Sobolev spaces of vector calculus and their transformation behavior. 
Let $\Omega \subseteq \bbR^{3}$ be a bounded open set. 
First, recall that $L^{p}(\Omega)$ is the space of scalar-valued $p$-integrable functions defined on $\Omega$
and that $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ for vector-valued functions with each component in $L^{p}(\Omega)$. 
In the three-dimensional setting, we are particularly interested in the Sobolev vector analysis. 
$W^p(\grad,\Omega) := W^{1,p}(\Omega)$ is the space of scalar-valued $L^{p}(\Omega)$ functions with weak gradients in $\bfL^{p}(\Omega)$, that is,  
\[
    W^p(\grad,\Omega) = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
\]
The space $\bfW^{p}(\curl,\Omega)$ of vector-valued $\bfL^{p}(\Omega)$ functions with weak curls in $\bfL^{p}(\Omega)$
and the space of vector-valued $\bfL^{p}(\Omega)$ functions with weak divergences in $L^{p}(\Omega)$ are defined as 
\[
    \bfW^{p}(\curl,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \curl \bfu \in \bfL^{p}(\Omega) \},
    \quad 
    \bfW^{p}(\divergence,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \divergence \bfu \in L^{p}(\Omega) \}.
\]
We are interested in transformations of these Sobolev tensor fields from one domain onto the other. 
Suppose that $\Omega, \Omega' \subset \mathbb{R}^3$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
The gradient-, curl-, divergence-conforming Piola transformations are the mappings 
$\phi^{\grad}: L^{p}(\Omega') \to L^{p}(\Omega)$,
$\phi^{\curl}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$, 
$\phi^{\divergence}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$,
and
$\phi^{\rm b}: L^{p}(\Omega') \to L^{p}(\Omega)$. 
They are defined by 
\begin{subequations}\label{eq_definition_piola}
\begin{align}
    \phi^{\grad}(v) &= v \circ \phi, \\
    \phi^{\curl}(\bfw) &= (\Jacobian\phi^{T} \bfw) \circ \phi, \\
    \phi^{\divergence}(\bfw) &= \left( \det(\Jacobian\phi) \Jacobian\phi^{-1} \bfw \right)  \circ \phi, \\
    \phi^{\rm b}(v) &= \left( \det\Jacobian\phi^{-1} v \right)  \circ \phi,
\end{align}
\end{subequations}
where $v \in L^{p}(\Omega')$ and $\bfw \in \bfL^{p}(\Omega')$, 
and where $\Jacobian\phi$ is the Jacobian matrix of $\phi$. 
These mappings are invertible. 
%
% TODO: bounds
%
We use the commutativity relations 
\begin{subequations}\label{eq_piola_commute}
\begin{align}
    \grad (\phi^{\grad} (v)) &= \phi^{\curl}(\grad v), 
    \\
    \curl (\phi^{\curl}(\bfv)) &= \phi^{\divergence}(\curl \bfv) 
    % TODO: watch the sign 
    \\
    \divergence (\phi^{\divergence}(\bfw)) &= \phi^{\rm b}(\divergence \bfw)
\end{align}
\end{subequations}
for $v \in W^{p}(\grad,\Omega)$, $\bfv \in \bfW^{p}(\curl,\Omega)$, and $\bfw \in \bfW^{p}(\divergence,\Omega)$.  
We summarize this as a commuting diagram:
\begin{align*}
    \begin{CD}
        W^{p}(\grad,\Omega') @>{\grad}>> \bfW^{p}(\curl,\Omega') @>{\curl}>> \bfW^{p}(\divergence,\Omega') @>{\divergence}>> L^{p}(\Omega')
        \\
        @VV{\phi^{\grad}}V 
        @VV{\phi^{\curl}}V 
        @VV{\phi^{\divergence}}V 
        @VV{\phi^{\rm b}}V 
        \\
        W^{p}(\grad,\Omega ) @>{\grad}>> \bfW^{p}(\curl,\Omega ) @>{\curl}>> \bfW^{p}(\divergence,\Omega ) @>{\divergence}>> L^{p}(\Omega )
    \end{CD}
\end{align*}

\begin{remark}
    The Piola transform goes int the opposite direction of the mapping $\phi : \Omega \rightarrow \Omega'$:
    scalar and vector fields over $\Omega'$ are transformed into scalar and vector fields over $\Omega$.
    This definition is in accordance with the notion of pullback, which we will review shortly. 
    One advantage of that definition is that it also makes sense whenever the transformation is not a diffeomorphism. 
    The literature also knows the Piola transform in the direction of the original mapping. 
\end{remark}

@@

Having discussed vector analysis, we review exterior calculus. 
We give a brief introduction into this topic, and refer to the literature for further background~\cite{...}. 
Let $V$ be a real vector space. 
Given an integer $k \geq 0$, we let $\Alt^{k}(V)$ denote the space of scalar-valued antisymmetric $k$-linear forms over $V$. 
Recall that any $k$-linear scalar-valued form $u$ over $V$ is called antisymmetric
if 
\[ 
    u( v_{\pi(1)}, v_{\pi(2)}, \ldots, v_{\pi(k)} ) 
    = 
    \text{sign}(\pi) 
    u( v_1, v_2, \ldots, v_k ) 
\]
for any $v_1, v_2, \dots, v_k \in V$ and any permutation $\pi$ of the indices \(\{1, 2, \ldots, k\}\). 
By definition, $\Alt^{1}(V)$ is just the dual space of $V$, and $\Alt^{0}(V)$ is the space of real numbers. 
Formally, we define $\Alt^{k}(V)$ to be the zero vector space when $k < 0$.

The wedge product (or exterior product) of alternating multilinear forms is a fundamental operation in exterior algebra. 
Given two alternating multilinear forms \( u_1 \in \Alt^{k}(V) \) and \( u_2 \in \Alt^{l}(V) \), 
their wedge product \( u_1 \wedge u_2 \) is a member of $\Alt^{k+l}(V)$
defined by the formula 
\[
    (u_1 \wedge u_2)(v_1, v_2, \ldots, v_{k+l}) 
    = 
    \frac{1}{k! l!} 
    % \sum_{ \pi \in S_{k+l}} 
    \sum_{ \pi } 
    \signum(\pi) 
    u_1( v_{\pi(  1)}, \ldots, v_{\pi(k  )} ) 
    u_2( v_{\pi(k+1)}, \ldots, v_{\pi(k+l)} ),
\]
for any \( v_1, v_2, \ldots, v_{k+l} \in V \).
Here, the sum runs over all permutations $\pi$ of the index set \(\{ 1, 2, \ldots, k+l \}\).
% Here, \( S_{k+l} \) denotes the symmetric group on \( k+l \) elements, and \( \signum(\pi) \) is the sign of the permutation \( \pi \). 
The exterior product is bilinear and associative, and satisfies 
\begin{align*}
    u_1 \wedge u_2 = (-1)^{kl} u_2 \wedge u_1,
    \quad 
    u_1 \in \Alt^{k}(V),
    \quad 
    u_2 \in \Alt^{l}(V).
\end{align*}
We are interested only in the special case $V = \bbR^{n}$ of alternating forms over the $n$-dimensional Euclidean space. 
Here, it is customary to identify $\Alt^{k}(V)$ with the space of antisymmetric tensors in $k$ indices. 
Moreover, this particular setting comes with a canonical basis. 
We let \(\{\cartanx^1, \cartanx^2, \ldots, \cartanx^{n}\}\) be the canonical basis dual to the standard unit vectors.
This is a canonical basis of $\Alt^{1}(\bbR^{n})$. 
To define a canonical basis of $\Alt^{k}(\bbR^{n})$, 
we first introduce $\Sigma(k,n)$, the set of strictly ascending mappings $\sigma : \{1,\dots,k\} \rightarrow \{1,\dots,n\}$,
and introduce the basic $k$-alternators 
\begin{align*}
    \cartanx^{\sigma} := \cartanx^{\sigma(1)} \wedge \dots \wedge \cartanx^{\sigma(k)}, \quad \sigma \in \Sigma(k,n). 
\end{align*}
These define a basis of $\Alt^{k}(\bbR^{n})$.
Note that 
$\dim \Alt^{k}(\bbR^{n}) = \binom{n}{k}$. 
In particular, $\Alt^{k}(\bbR^{n})$ is the zero vector space whenever $k > n$.
% The dimension of the space of \(k\)-forms, denoted \(\bigwedge^k V\), is given by the binomial coefficient \(\binom{n}{k}\).

@@

We write $C^{\infty}\Alt^{k}(\Omega)$ for the space of smooth differential $k$-forms over $\Omega \subseteq \bbR^{n}$,
which is the vector space of smooth mappings from $\Omega$ into $\Alt^{k}(\bbR^{n})$.
The exterior derivative \( \cartan \) is an operator that takes a \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) 
to a \((k+1)\)-form \( \cartan\omega \in C^{\infty}\Alt^{k+1}(\Omega) \). 
Every \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) can be written 
\[
    \omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \omega_{\sigma} \, 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)},
\]
where \( \omega_{\sigma} : \Omega \rightarrow \bbR \) are smooth functions.
The exterior derivative \( \cartan\omega \) is defined by
\[
    \cartan\omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \sum_{j=1}^{n} 
    \frac{\partial \omega_{\sigma}}{\partial x^j} 
    \, \cartanx^j \wedge 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)}.
\]
The exterior derivative is linear and nilpotent, which means 
\( \cartan(\cartan\omega) = 0 \) for any \( \omega \in C^{\infty}\Alt^{k}(\Omega) \).
Moreover, it satisfies the Leibniz rule:
\[ 
    \cartan(\omega \wedge \eta) 
    = 
    \cartan\omega \wedge \eta + (-1)^{k} \omega \wedge \cartan\eta, 
    \qquad \omega \in C^{\infty}\Alt^{k}(\Omega), 
    \qquad \eta \in C^{\infty}\Alt^{l}(\Omega).
\]
Let us now turn our attention to Sobolev spaces of differential forms. 
Since $\Alt^{k}(\Omega)$ is a normed space, we can consider the pointwise norms of differential $k$-forms. 
% First, we notice that the % TODO: introduce a norm 
We let $L^{p}\Alt^{k}(\Omega)$ be the space of differential $k$-forms over $\Omega$ with locally integrable coefficients 
such that its pointwise norm is $p$-integrable. 
The exterior derivative is defined in the sense of distributions and we introduce 
\[
    W^p\Alt^{k}(\Omega) = \{ u \in L^{p}\Alt^{k}(\Omega) \suchthat \cartan v \in L^{p}\Alt^{k+1}(\Omega) \}
\]
We are interested in transformations of these Sobolev tensor fields from one domain onto the other. 
Suppose that $\Omega,\Omega' \subset \bbR^{n}$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
The pullback of $u \in W^p\Alt^{k}(\Omega')$ along $\phi$ is the differential form 
\[ 
    \phi^{\ast}u_{|x}( v_{1}, v_{2}, \ldots, v_{k} ) 
    = 
    u_{|\phi(x)}( \Jacobian\phi_{|x} \cdot v_1, \Jacobian\phi_{|x} \cdot v_2, \ldots, \Jacobian\phi_{|x} \cdot v_k ) 
\]
for any $v_1, v_2, \dots, v_k \in \bbR^{n}$ and any $x \in \Omega$. One can show that $\phi^{\ast}u \in W^p\Alt^{k}(\Omega)$.

% TODO: transformationssatz

\begin{remark}
    In three dimensions, 
    the calculus of differential forms can be translated into classical vector calculus. 
    This is expressed formally as the commuting diagram 
    \[
    \begin{CD}
        C^\infty\Alt^0(\Omega) @>{\cartan}>> C^\infty\Alt^1(\Omega) @>{\cartan}>> C^\infty\Alt^2(\Omega) @>{\cartan}>> C^\infty\Alt^3(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        @VV{\varpi^{3}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{3} @>{\curl}>> C^\infty(\Omega)^{3} @>{\divergence}>> C^\infty(\Omega)
    \end{CD}
    \]
    where $\varpi^{0}$ and $\varpi^{3}$ are the identity mappings and where 
    \begin{align*}
     \varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} + u_3 \cartanx^{3} \right) 
     &= 
     \left( u_1, u_2, u_3 \right), 
     \\
     \varpi^{2}\left( u_{12} \cartanx^{1} \wedge \cartanx^{2} + u_{13} \cartanx^{1} \wedge \cartanx^{3} + u_{23} \cartanx^{2} \wedge \cartanx^{3} \right) 
     &= 
     \left( u_{23}, - u_{13}, u_{12} \right).   
    \end{align*}
    In two dimensions, 
    the calculus of differential forms can be translated into 2D vector calculus in two different ways. 
    To the authors' best knowledge, neither convention is dominant in the literature.
    We summarize the situation in the following commuting diagram: 
    \[ % TODO: what is the convention for the curl and rotation ? look up in the textbooks?
    \begin{CD}
        C^\infty(\Omega) @>{\curl}>> C^\infty(\Omega)^{2} @>{\divergence}>> C^\infty(\Omega)
        \\
        @AA{\varkappa^{0}}A 
        @AA{\varkappa^{1}}A 
        @AA{\varkappa^{2}}A 
        \\
        C^\infty\Alt^0(\Omega) @>{\cartan}>> C^\infty\Alt^1(\Omega) @>{\cartan}>> C^\infty\Alt^2(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{2} @>{\rot}>> C^\infty(\Omega)
    \end{CD}
    \]
    Here, $\varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} \right) = \left( u_1, u_2 \right)$ is the lower middle isomorphism. We introduce the rotation operator $J(x,y) = (y,-x)$ and define $\varkappa = J \varpi$ and $\rot = \divergence J$.
    The other vertical arrows are the identity. 
    The utility of exterior calculus is that the operators of vector calculus can be translated into a common framework that does not depend on the dimension.
\end{remark}

% TODO: spaces with boundary conditions: grad, curl, divergence, exterior derivative; approximation properties

@@

@@

    \cartan \Bogov_{\ell,a} u(x) 

@@

@@

    \Bogov_{\ell+1,a} \cartan u(x)

@@

In particular, if $\cartan u = 0$, then $\cartan \Bogov_{\ell,a} u = u$.
